---
title: 'Problem Set #4'
author: 'Gaurav Khanna'
output:
  pdf_document: default
urlcolor: blue
---

<!--
Some guidelines for submitting problem sets in this course:

- Please submit a PDF document rather than a Word document or a Google document.
- Please put your name at the top of your problem set.
- Please **bold** or *highlight* your numerical answers to make them easier to find.
- If you'll be using `R` or `Python` code to calculate your answers, please put the code and its output directly into your Problem Set PDF document.
- It is highly recommended, although not required, that you use the RMarkdown feature in RStudio to compose your problem set answers. RMarkdown allows you to easily intermingle analysis code and answers in one document. It is of a similar design as `jupyter` and an ipython notebook.
- You do not need to show work for trivial calculations, but showing work is always allowed.
- For answers that involve a narrative response, please feel free to describe the key concept directly and briefly, if you can do so, and do not feel pressure to go on at length.
- Please ask us questions about the problem set if you get stuck. **Don’t spend more than 20 minutes puzzling over what a problem means.** 
- Please ensure that someone (us!) can compile your solution set. The best way is to use the web-hosted links we've provided. 
-->

```{r}
# load packages 
library(foreign)
library(data.table)
library(dplyr)
library(experiment)
```
***
# 1. Potential Outcomes 
## a. Make up a hypothetical schedule of potential outcomes for three Compliers and three Never-Takers where the ATE is positive but the CACE is negative. By ATE, we mean the average treatment effect for the entire population, including both compliers and never-takers.  Note that we can never compute this ATE directly in practice, because we never observe both potential outcomes for any individual, especially for never-takers. That's why this question requires you to provide a complete table of hypothetical potential outcomes for all six subjects.

```{r}
# Simulating the experiment
# Hypothetical schedule of all the observations

# Control
dz0 <- c(rep(0, 6))
# Treatment, Compliars and Never Takers
dz1 <- c(rep(0,3), rep(1, 3))
# Output variable in control
yd0 <- c(1, 3, 7, 2, 1, 6)
# Output variable in treatment
yd1 <- c(4, 3, 7, 3, 2, 2)

# Converting to a data frame
exp.df <- data.frame(dz0, dz1, yd0, yd1)
# TE for each
exp.df$te <- exp.df$yd1 - exp.df$yd0
exp.df

```

**They hypothetical schedule is above**

```{r}
# Calculations for ATE & CACE

# ATE
est.ate <- mean(exp.df$yd1) - mean(exp.df$yd0)
print('ATE')
est.ate

# Subset DF for compliars
exp.comp <- exp.df[which(dz1==1),]
print('DF of compliars')
exp.comp

# Since we have the entire schedule CACE is just the average treatment effect for the compliars
# We'll use the subset of the compliars created above
exp.cace <- mean(exp.comp$yd1 - exp.comp$yd0)
print('CACE')
exp.cace

```


**Answer: If we use the entire schedule (hypothetical) we get an ATE of .166 and the CACE of -.667**

###Calculating the ATE estimate without the entire schedule (like in an expreiment, by calculating ATE and then CACE by dividing with the ratio of compliars)
```{r}
# The ratio of compliars is 50%
exp.cace1 = est.ate/.5
exp.cace1
```
## b. Suppose that an experiment were conducted on your pool of subjects. In what ways would the estimated CACE be informative or misleading? 

**TBD: Answer: In an experiment, we would not have the full schedule of observations. We'll be able to calculate the ATE and then CACE (cace1) by using ATE and the ratio of compliars. Calculating the CACE that way gives us an estimate of .333 which is misleading. **

## c. Which population is more relevant to study for future decision making: the set of Compliers, or the set of Compliers plus Never-Takers? Why?

**Answer: The population of compliers is more important for future decision making. That's where the information on real effect of the treatment comes from.An example would be trying to measure the efficacy of a medicine/pill**
*There may be some reasons to briefly look at "never takers". It is to get a idea of the ratio of the 2 groups in the overall population. This may be useful in some experiments where we are interested in the "intent to treat" effect instead of the "actual" treatment effect on the treated. Many expriments to influence policy (like rebates for buying health insurance) fall in this category*

***

# Turnout to Vote 
Suppose that a researcher hires a group of canvassers to contact a set of 1,000 voters randomly assigned to a treatment group. When the canvassing effort concludes, the canvassers report that they successfully contacted 500 voters in the treatment group, but the truth is that they only contacted 250. When voter turnout rates are tabulated for the treatment and control groups, it turns out that 400 of the 1,000 subjects in the treatment group voted, as compared to 700 of the 2,000 subjects in the control group (none of whom were contacted). 

## a. If you believed that 500 subjects were actually contacted, what would your estimate of the CACE be? 
```{r}
# Overall turnout rate in the control group
turn.control <- 700/2000
turn.control
# Overall turnout rate in the treatment group
turn.treat <- 400/1000
turn.treat

itt <- turn.treat - turn.control
itt

```
```{r}
# CACE on assumption that 500 subjects were contacted
cace1 <- itt/(500/1000)
cace1
```
**Answer: CACE would be .1**

## b. Suppose you learned that only 250 subjects were actually treated. What would your estimate of the CACE be? 
```{r}
# CACE on assumption that 250 subjects were contacted
cace2 <- itt/(250/1000)
cace2
```
**Answer: CACE would be .2 in this case**

## c. Do the canvassers' exaggerated reports make their efforts seem more or less effective? Define effectiveness either in terms of the ITT or CACE. Why does the definition matter? 
**Answer: The exaggerated reports make the canvassers' efforts seem less effective. We get a lower estimate for CACE with the exaggerated reports**
**Effectiveness** can be defined in terms of either ITT or CACE. We could either define it as follows:
1. +ve value (greater than 0) for the ITT
2. +ve value for CACE, greater than ITT
**The definition matters as it forces us to recognize exactly what we want to measure before the expriment.**
**In this specific example, we'll concentrate on ITT if we're trying to measure the effectiveness of the intent, i.e. knocking on doors**
**We'll concentrate on CACE if our concern is limited to measure the impact of actually talking to people who open the door**
**Defining the primary outcome before the experiment would help pick the right metric for measurement**

*** 

# 3. Turnout in Dorms
Guan and Green report the results of a canvassing experiment conduced in Beijing on the eve of a local election. Students on the campus of Peking University were randomly assigned to treatment or control groups. Canvassers attempted to contact students in their dorm rooms and encourage them to vote. No contact with the control group was attempted. Of the 2,688 students assigned to the treatment group, 2,380 were contacted. A total of 2,152 students in the treatment group voted; of the 1,334 students assigned to the control group, 892 voted. One aspect of this experiment threatens to violate the exclusion restriction. At every dorm room they visited, even those where no one answered, canvassers left a leaflet encouraging students to vote. 

```{r}
library(foreign)
d3 <- read.dta("./data/Guan_Green_CPS_2006.dta")
head(d3, 20)
```
```{r}
# Finding more about the data
nrow(d3)
summary(d3)
```
```{r}
# plot of "treatment" for dorm ID's
plot(d3$dormid, d3$treat2)
```

## a. Using the data set from the book's website, estimate the ITT. First, estimate the ITT using the difference in two-group means. Then, estimate the ITT using a linear regression on the appropriate subset of data. *Heads up: There are two NAs in the data frame. Just na.omit to remove these rows.*

```{r}

# Getting key values out of the data

# Number in treated
exp3.treat.n = 2688
# Number that voted from treatment
exp3.treat.v = 2152
# Number that were contacted
exp3.treat.c = 2380

# Number in control
exp3.cont.n = 1334
# Number that voted
exp3.cont.v = 892

# Conversion ratio in treatment
exp3.treat.convert = exp3.treat.v/exp3.treat.n
paste0("Conversion ratio in treatment: ", exp3.treat.convert)

# Conversion ratio in control
exp3.cont.convert = exp3.cont.v/exp3.cont.n
paste0("Conversion ratio in control: ", exp3.cont.convert)

# Ratio of actually treated
exp3.treat.ar = exp3.treat.c/exp3.treat.n
paste0("Ratio of actually treated or alpha: ", exp3.treat.ar)

```

### ITT using difference in 2 group means. 
```{r}
itt.gm = exp3.treat.convert - exp3.cont.convert
paste0("ITT using difference in means: ", itt.gm)
```
### ITT using linear regression
```{r}
# Linear regression
lr3 <- lm(turnout ~ treat2, data=d3, na.action = na.omit)
summary(lr3)
paste0("ITT with regression: ", lr3$coefficients['treat2'])
```

**Answer**
**ITT from difference in means: 0.132**
**ITT from regression: 0.132 (0.01422)**


## b. Use randomization inference to test the sharp null hypothesis that the ITT is zero for all observations, taking into account the fact that random assignment was clustered by dorm room. Interpret your results. 

```{r}
# Pull out the cluster ID's
clusters.all <- unique(d3$dormid)
# How many clusters
clusters.n = length(all_clusters)
clusters.n
```
```{r}
head(clusters.all)
```


```{r}
# Generic function to pick the cluster ID's for treatment

clusters.in.treatment = clusters.n/2
paste0("Clusters in treatment: ", clusters.in.treatment)

randomize.clusters <- function(d) {
  treat.clusters <- sample(x = clusters.all, 
                                size = clusters.in.treatment, 
                                replace = FALSE)
  # returns 1 if a cluster ID in those that are picked
  return(as.numeric(d$dormid %in% treat.clusters))
}

# Generic function to estimate the ATE
est.ate <- function(result, treat) { 
  mean(result[treat==1]) - mean(result[treat==0])
  } 
```

```{r}
# Trying out one randomization. Now we have the data prepared for the experiment
# First creating a new DF
d31 <- read.dta("./data/Guan_Green_CPS_2006.dta")
d31$treat2 <- randomize.clusters(d31)

```

```{r}
# Finding out more about the new data set
clusters <- group_by(d31, dormid)
clusters.summary <- summarize(clusters, count=n(), mean_y=mean(turnout), mean_treat2=mean(treat2) )
clusters.summary
```
```{r}
# Validating the data
mean(d31$turnout[d31$treat2 == 1])
mean(d31$turnout[d31$treat2 == 0])

# Got to remove the Rows with na
nrow(d31)
d31 <- na.omit(d31)
nrow(d31)


# Now validate
mean(d31$turnout[d31$treat2 == 1])
mean(d31$turnout[d31$treat2 == 0])
```

```{r}
# ATE for one randomization
ateR <- est.ate(d31$turnout, d31$treat2)
ateR
```
```{r}
# 1K Randomizations
# Repeating the expriment 10,000 times
distribution.under.sharp.null <- replicate(1000, est.ate(d31$turnout, (randomize.clusters(d31))))
ate.mean <- mean(distribution.under.sharp.null)
paste0("Mean ITT from 5k randomizations: ", ate.mean)

```
```{r}
# Graph for the estimates
plot(density(distribution.under.sharp.null), 
     main = "Density under Sharp Null")
abline(v = itt.gm, col = "blue")
```
**TODO**

## c. Assume that the leaflet had no effect on turnout. Estimate the CACE. Do this in two ways: First, estimate the CACE using means. Second, use some form of linear model to estimate this as well. If you use a 2SLS, then report the standard errors and draw inference about whether the leaflet had any causal effect among compliers. 

### CACE using means
```{r}
# CACE using the means. ITT/alpha
cace.m <- itt.gm/exp3.treat.ar
paste0('CACE using means: ', cace.m)

```
### CACE using a linear model

We'll split treatment into 2 variables for the new linear model:
1. "contact": This already exists in the data. Its when the door is answered
2. "leaflet": Generated independent variable (treat2 * !contact). Its when the door is not answered and a leaflet is dropped
Our model tries to measure the effect of contact and leaflet separately using the regression below

```{r}
# Adding a new variable "leaflet"
d3$leaflet <- d3$treat2 * !d3$contact
head(d3, 20)
summary(d3)
```

```{r}
lr33 <- lm(turnout ~ contact + leaflet, data=d3, na.action = na.omit)
summary(lr33)
# paste0("ITT with regression: ", lr33$coefficients['treat2', 'contact'])
```
**Answer: CACE using means is 0.14900280951961**  
**The linear model for estimating CACE is as follows: **  
**turnout = intercept + a(Contact) + b(leaflet) + errors**  
**We try to find out the impact of contact and leaflet (when there's no contact but a leaflet is dropped) separately**  
**The coefficients from the model are as follows: **  
Coefficients:
            Estimate Std. Error t value Pr(>|t|)    
(Intercept)  0.66867    0.01162  57.526  < 2e-16 ***
contact      0.13562    0.01452   9.341  < 2e-16 ***
leaflet      0.10332    0.02687   3.845 0.000123 ***

**We can interpret the coefficient of contact as the CACE: .136 (.0145)**  
**The model shows that dropping the leaflet as a significant positive effect on the outcome 0.10332 (0.02687).**

***

# 4. Why run a placebo? 
Nickerson describes a voter mobilization experiment in which subjects were randomly assigned to one of three conditions: a baseline group (no contact was attempted); a treatment group (canvassers attempted to deliver an encouragement to vote); and a placebo group (canvassers attempted to deliver an encouragement to recycle). Based on the results in the table below answer the following questions 

+----------------------+-----------+------+---------+
| Treatment Assignment | Treated ? | N    | Turnout |
+======================+===========+======+=========+
| Baseline              | No       | 2572 | 31.22%  |
+----------------------+-----------+------+---------+
| Treatment            | Yes       | 486  | 39.09%  |
+----------------------+-----------+------+---------+
| Treatment            | No        | 2086 | 32.74%  |
+----------------------+-----------+------+---------+
| Placebo              | Yes       | 470  | 29.79%  |
+----------------------+-----------+------+---------+
| Placebo              | No        | 2109 | 32.15%  |
+----------------------+-----------+------+---------+
 
**First** Use the information to make a table that has a full recovery of this data. That is, make a `data.frame` or a `data.table` that will have as many rows a there are observations in this data, and that would fully reproduce the table above. (*Yes, this might seem a little trivial, but this is the sort of "data thinking" that we think is important.*)

a. Estimate the proportion of Compliers by using the data on the Treatment group.  Then compute a second estimate of the proportion of Compliers by using the data on the Placebo group.  Are these sample proportions statistically significantly different from each other?  Explain why you would not expect them to be different, given the experimental design. (Hint: ITT_D means "the average effect of the treatment on the dosage of the treatment." I.E., it’s the contact rate $\alpha$ in the async).

b. Do the data suggest that Never Takers in the treatment and placebo groups have the same rate of turnout? Is this comparison informative? 

c. Estimate the CACE of receiving the placebo. Is this estimate consistent with the substantive assumption that the placebo has no effect on turnout? 

d. Estimate the CACE of receiving the treatment using two different methods. First, use the conventional method of dividing the ITT by the ITT_{D}. (This should be a treatment vs. control comparison.)

```{r}

```

e. Then, second, compare the turnout rates among the Compliers in both the treatment and placebo groups. Interpret the results. 

```{r}

```

f. Based on what we talked about in class -- that the rate of compliance determines whether one or another design is more efficient -- given the compliance rate in this study, which design *should* provide a more efficient estimate of the treatment effect? If you want to review the specific paper that makes this claim, check out [this link](https://github.com/UCB-MIDS/experiments-causality/blob/master/readings/GerberGreenKaplanKern.2010.pdf). Does it? 


<!--
# EVERYTHING IN THIS COMMENTED SECTION IS NOT REQUIRED. THESE ARE GOOD PROBLEMS, AND IF YOU WANT TO CHECK YOUR 
# UNDERSTANDING, THEY WOULD BE GOOD TO DO. 

# More Practice 
Determine the direction of bias in estimating the ATE for each of the following situations when we randomize at the individual level.  Do we over-estimate, or underestimate? Briefly but clearly explain your reasoning.

a. In the advertising example of Lewis and Reiley (2014), assume some treatment-group members are friends with control-group members.

b. Consider the police displacement example from the bulleted list in the introduction to FE 8, where we are estimating the effects of enforcement on crime.

c. Suppose employees work harder when you experimentally give them compensation that is more generous than they expected, that people feel resentful (and therefore work less hard) when they learn that their compensation is less than others, and that some treatment-group members talk to control group members.

d. When Olken (2007) randomly audits local Indonesian governments for evidence of corruption, suppose control-group governments learn that treatment-group governments are being randomly audited and assume they are likely to get audited too.


# FE exercise 8.2
National surveys indicate that college roommates tend to have correlated weight. The more one roommate weights at the end of the freshman year, the more the other freshman roommate weights. On the other hand, researchers studying housing arrangements in which roommates are randomly paired together find no correlation between two roommates' weights at the end of their freshman year. *Explain how these two facts can be reconciled.*
-->


# 5. Tetris FTW? 
A doctoral student conducted an experiment in which she randomly varied whether she ran or walked 40 minutes each morning. In the middle of the afternoon over a period of 26 days she measured the following outcome variables: (1) her weight; (2) her score in Tetris; (3) her mood on a 0-5 scale; (4) her energy; and (5) whether she got a question right on the math GRE. 

```{r}
d <- read.dta("./data/Hough_WorkingPaper_2010.dta")
head(d)
``` 

a. Suppose you were seeking to estimate the average effect of running on her Tetris score. Explain the assumptions needed to identify this causal effect based on this within-subjects design. Are these assumptions plausible in this case? What special concerns arise due to the fact that the subject was conducting the study, undergoing the treatments, and measuring her own outcomes? 

b. Estimate the effect of running today on Tetris score. What is the ATE?


```{r}

```

c. One way to lend credibility to with-subjects results is to verify the no-anticipation assumption. Construct a regression using the variable `run` to predict the `tetris` score *on the preceding day*. Presume that the randomization is fixed. Why is this a test of the no-anticipation assumption? Does a test for no-anticipation confirm this assumption? 

```{r}

```

d. Now let's use regression to put a standard error on our ATE estimate from part (b). Regress Tetris score on the the variable `run`, this time using the current rather than the future value of `run`.  Is the impact on Tetris score statistically significant? 

e. If Tetris responds to exercise, one might suppose that energy levels and GRE scores would as well. Are these hypotheses borne out by the data?  

f. Suppose the student decides to publish her results on Tetris, since she finds those most interesting.  In the paper she writes, she chooses to be concise by ignoring the data she collected on energy levels and GRE scores, since she finds those results less interesting.  How might you criticize the student's decision?  What trap may she have fallen into?

g. After submitting her paper to a journal, the student thinks of another hypothesis.  What if running has a relatively long-lasting effect on Tetris scores?  Perhaps both today's running and yesterday's running will affect Tetris scores.  Run a regression of today's Tetris score on both today's `run` variable and yesterday's `run` variable.  How does your coefficient on running today compare with what you found in part (d)?  How do you interpret this comparison?

h. (optional) Note that the observations in our regression are not necessarily independent of each other. An individual might have serially correlated outcomes, regardless of treatment.  For example, I might find that my mood is better on weekends than on weekdays, or I might find that I'm terrible at playing Tetris in the few days before a paper is due, but I get better at the game once my stress level has lowered. In computing standard errors for a regression, OLS assumes that the observations are all independent of each other.  If they are positively serially correlated, it's possible that OLS will underestimate the standard errors.

To check this, let's do randomization inference in the regression context.  Recall that the idea of randomization inference is that under the sharp null hypothesis, we can re-randomize, recompute the ATE, and get approximately the right answer (zero) for the treatment effect.  So, returning to the regression we ran in part (g), please generate 1000 new randomizations of the `run` variable, use those to replace the current and lagged values of `run` in your dataset, then run the regression again.  Record the coefficient you get on the contemporaneous value of `run`, and repeat this re-randomization exercise 1000 times.  Plot the distribution of beta. What are the 2.5% and 97.5% quantiles?  How do they compare with the width of the 95% confidence interval you got for your main `run` coefficient in the regression in part (g)?
